Stack Safety as a Security Property

### 2020-06-03 main.tex
Moved updates into main.tex, with a few other changes to preliminaries. Focusing on properties next.

    {\it Eager stack confidentiality} states that data outside a callee's stack frame is secret,
    and does not interfere with its observable behavior or with the machine state at return. Its
    weaker variant {\it observable stack safety} lifts the latter restriction and states that
    the callee may not leak these secrets directly or by changing the behavior of its caller.

@@

\leo{I tried the ? notation for the potentially empty trace, which reveals that
  we're slightly informal in the second case. Expand?}
We define the append operation \(T \cdot T?\), where \(T?\) may be a trace or empty (\(\varepsilon\)):
\[T \cdot T? \triangleq
  \begin{cases}
    T & \text{if } T? = \varepsilon \\
    t T? & \text{if } T = t \\
    t T' \cdot T? & \text{if } T = t T' \\
  \end{cases}\]

We define the ``trace-of'' operator \(\hookrightarrow\) that coinductively
relates a machine state with the trace of machine states paired with
observations produced by repeated application of step:
%
\[\frac{\mach_0 \stepstoobs{\obs} \mach_1 \hspace{.5cm} \mach_1 \hookrightarrow \machT}
       {\mach_0 \hookrightarrow (\mach_0,\obs) \machT}\]
%
Since the step function is total, if \(\mach \hookrightarrow \machT\),
then \(\machT\) must be infinite.

We will also wish to take a prefix of the trace from an initial state
up through the first state where some condition holds. Let \(f\) be
such a predicate on machine states.  Then we define \(\mach
\hookrightarrow \machT | f\) (read ``\(\machT\) is the prefix from
\(\mach\) until \(f\)''):\leo{Abrupt. Explain in words?}
    \[\frac{\mach_0 \stepstoobs{\obs} \mach_1 \hspace{.5cm} \neg f\ \mach_1 \hspace{.5cm} \mach_1
              \hookrightarrow \machT | f}
           {\mach_0 \hookrightarrow (\mach_0,\obs) \machT | f}\]
    \[\frac{f\ m}
           {m \hookrightarrow (m,\tau)}\]

Similarly, we define \(\hookrightarrow_P\) to relate a machine-policy
state pair to the trace of triples of machine states, policy states,
and observations induced by by \(\stepstoobs{}_P\):
%
\leo{I think this would read better as not exists. Thoughts?}
\[\frac{\forall \pol_1 . \pol_0 \not \longrightarrow \pol_1}
       {\mpstate{\mach_0}{\pol_0} \hookrightarrow \mpostate{\mach_0}{\pol_0}{\tau}}\]
\[\frac{\mach_0 \stepstoobs{\obs} \mach_1 \hspace{.5cm} \pol_0 \longrightarrow \pol_1 \hspace{.5cm}
  \mpstate{\mach_1}{\pol_1} \hookrightarrow \machT}
       {\mpstate{\mach_0}{\pol_0} \hookrightarrow_P \mpostate{\mach_0}{\pol_0}{\obs} \machT}\]

We also lift the prefix-up-to relation to such traces.
\[\frac{\mpstate{\mach_0}{\pol_0} \xrightarrow{\obs} \mpstate{\mach_1}{\pol_1} \hspace{.5cm}
  \neg f\ \mach_1 \hspace{.5cm} \mpstate{\mach_1}{\pol_1} \hookrightarrow_P \machT | f }
       {\mpstate{\mach_0}{\pol_0} \hookrightarrow_P \mpostate{\mach_0}{\pol_0}{\obs} \machT | f}\]

\[\frac{\mpstate{\mach_0}{\pol_0} \xrightarrow{\obs} \mpstate{\mach_1}{\pol_1} \hspace{.5cm} f\ \mach_1}
       {\mpstate{\mach_0}{\pol_0} \hookrightarrow_P \mpostate{\mach_0}{\pol_0}{\obs}
         \mpostate{\mach_1}{\pol_1}{\tau} | f}\]

@@

Let the predicate \(\isinf(T)\) hold when the trace $T$ is infinite.
We then define {\em policy-sensitive observational equivalence}: a
machine-policy-observation\leo{Call these MPO traces?} is equivalent
to a machine-observation trace, if the MPO trace is infinite and the
observations of MPO and M coincide, or if the MPO trace is finite
(e.g. due to a policy fault) and its observations are a prefix of
those of M.

\leo{Why $\mathit{Obs}$ and not $\pi_o$ defined earlier?}
\[\frac{\neg \isinf(\MPT) \hspace{.5cm} \mathit{Obs}(\MPT) \lesssim \mathit{Obs}(\machT)}
           {\MPT \sim_P \machT}\]

\[\frac{\isinf(\MPT) \hspace{.5cm} \mathit{Obs}(\MPT) \simeq \mathit{Obs}(\machT)}
           {\MPT \sim_P \machT}\]

>>>>

    Confidentiality is in the style of noninterference, defined on a per-call basis. For each call,
    the caller's data and the uninitialized stack do not interfere with observable behavior until return.
    In {\it eager stack confidentiality} this means that when a function is called, the memory outside of
    its stack frame could be scrambled without changing its observable behavior or the changes it makes
    to the program state. This implies the weaker {\it observable stack confidentialty} in which
    inaccessible memory could be scrambled at each call and restored on return without changing the
    observable behavior of the remaining program.

@@

We define a few operations on traces generically.
The {\it split} operation \(\mathit{split}(f,T)\) takes a trace \(T\) and a predicate on elements \(f\) and
separates \(T\) into a prefix and suffix at the first element on which \(f\) holds:
  \[\frac{f\ t}
         {\mathit{split}(f,tT) = t, tT}\]
  \[\frac{\neg f\ t \hspace{.5cm} \mathit{split}(f,T) = T_1,T_2}
         {\mathit{split}(f,tT) = tT_1,T_2}\]

The {\it join} operation \(T_1 \cdot T_2\) takes two traces and links them by replacing the last element
of the first (if it has one) with the first of the second:
  \[\frac{}
         {T_1t \cdot T_2 = T_1 T_2}\]

We define the ``trace-of'' operator \(\hookrightarrow\) that coinductively
relates a machine state with the trace of machine states paired with
observations produced by repeated application of step:
\[\machT \in \mathit{MOTrace} = \mathit{Trace}\ (\MACHS \times \OBSS)\]
\[\hookrightarrow : \machT \rightarrow \mathit{MOTrace}\]
%
\[\frac{\mach_0 \stepstoobs{\obs} \mach_1 \hspace{.5cm} \mach_1 \hookrightarrow \machT}
       {\mach_0 \hookrightarrow (\mach_0,\obs) \machT}\]
%
Since the step function is total, if \(\mach \hookrightarrow \machT\),
then \(\machT\) must be infinite.

\sna{This is a bit off, because \(f\) is a predicate on machine states, not the pairs (and later triples).
     Will think about what the best way is to present this.}
We will also wish to take a prefix of the trace from an initial state up through the first state where
some condition holds, using {\it split}. Let \(f\) be a predicate on machine states. Then we define
\(\mach \hookrightarrow \machT | f\) (read ``\(\machT'\) is the prefix of \(\machT\) up to \(f\)''):
  \[\frac{\mach \hookrightarrow \machT \hspace{.5cm} \mathit{split}(f,\machT) = \machT_1,\machT_2}
         {\mach \hookrightarrow \machT_1 | f}\]

Similarly, we define \(\hookrightarrow_P\) to relate a machine-policy
state pair to the trace of triples of machine states, policy states,
and observations induced by by \(\stepstoobs{}_P\):
%
\leo{I think this would read better as not exists. Thoughts?}
\sna{Looks good.}
  \[\MPT \in \mathit{MPOTrace} = \mathit{Trace}\ (\MACHS \times \POLS \times \OBSS)\]
  \[\hookrightarrow_P : \MACHS \times \POLS \rightarrow \mathit{MPOTrace}\]
  \[\frac{\neg \exists \pol_1 . \pol_0 \rightharpoonup \pol_1}
         {\mpstate{\mach_0}{\pol_0} \hookrightarrow \mpostate{\mach_0}{\pol_0}{\tau}}\]
  \[\frac{\mach_0 \stepstoobs{\obs} \mach_1 \hspace{.5cm} \pol_0 \rightharpoonup \pol_1 \hspace{.5cm}
            \mpstate{\mach_1}{\pol_1} \hookrightarrow \machT}
         {\mpstate{\mach_0}{\pol_0} \hookrightarrow_P \mpostate{\mach_0}{\pol_0}{\obs} \machT}\]

We also lift the prefix-up-to relation to such traces.
  \[\frac{\mpstatename \hookrightarrow \MPT \hspace{.5cm} \mathit{split}(f,\MPT) = \MPT_1,\MPT_2}
         {\mpstatename \hookrightarrow \MPT_1 | f}\]

@@

Let the predicate \(\isinf(T)\) hold when the trace $T$ is infinite.
We then define {\em policy-sensitive observational equivalence}: an
{\it MPOTrace} \(\MPT\) is equivalent to an {\it MOTrace} trace \(\machT\) if \(\MPT\)
is infinite and the observations of \(\MPT\) and \(\machT\) coincide, or if the \(\MPT\)
is finite (due to a policy fault) and its observations are a prefix of those of M.

% Fixed by Sean
%\leo{Why $\mathit{Obs}$ and not $\pi_o$ defined earlier?}
\[\frac{\neg \isinf(\MPT) \hspace{.5cm} \pi_o(\MPT) \lesssim \pi_o(\machT)}
           {\MPT \sim_P \machT}\]

\[\frac{\isinf(\MPT) \hspace{.5cm} \pi_o(\MPT) \simeq \pi_o(\machT)}
           {\MPT \sim_P \machT}\]
